---
title: "SARS-CoV-2 transmission in students of public schools of Catalonia, Spain, after a month of reopening"
author: "Llupia et al"
date: "Dec. 22nd 2020"
output:
  html_document: null
  pdf_document: default
  toc_depth: 2
  toc: yes
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
library(lubridate)
library(incidence)
library(ggplot2)
library(lme4)
library(knitr)
library(epitools)
library(sjPlot)
```
## Declaration

**This is supporting information alongside the article "SARS-CoV-2 transmission in students of public schools of Catalonia, Spain, after a month of reopening" by Llupià et al.**

## Dataset description

Read the dataset 'total.centres.csv' where all the data for the analysis is stored:
```{r}
total.centres<-read.csv('total-centres.csv')

#Generate smaller database of Primary Education
primary.schools<-subset(total.centres,total.centres$Type.of.School=="Public Primary 3-12")
#Generate smaller database of Secondary Education
secondary.schools<-subset(total.centres,total.centres$Type.of.School=="Public Secondary 12-16+")
```

All data comes from publicly available data of the Catalan Government (see below 'Data Sources'). For convenience, the dataset 'total-centres.csv' contains the all the data for the analysis. Every line is a different centre registered in the `Traça Covid' system for reporting Sars-Cov-2 cases among students and staff (see the paper for description).

The variables are the following:


* 'Students.Weeks.4.and.5' : Incidence among students in weeks 4 and 5,
* 'TOTAL.Weeks.1.and.2' : Reported Cases among students and staff in weeks 1 and 2.
* 'Students.Weeks.1.and.2' : Reported Cases among students in weeks 1 and 2,
* 'Adults.Weeks.1.and.2' :Reported Cases among staff in weeks 1 and 2,
* 'BHA.Incidence.Weeks.1.and.2' : Incidence in the Basic Health Area (cases per 100,000 habitants per 14 days) in weeks 1 and 2,
* 'Type.of.School' : Type of School, primary or Secondary.
* 'Educational.Levels' : Type of studies in the centre (used to determine age range).
* 'Confined.Students.Week.2' : Number of confined students at the end of week 2.       
* 'Confined.Students.Week.3' : Number of confined students at the end of week 3.
* 'Confined.Students.Week.5' : Number of confined students at the end of week 5.


## Data Sources

* The database of ED centres, from the Education department: [https://analisi.transparenciacatalunya.cat/d/3u9c-b74b]. Includes the type of studies. 
* `Enrolled.Students` The number of enrolled students in the previous course 2019/20 by ED centre. [https://analisi.transparenciacatalunya.cat/Educaci-/Alumnes-Enrolled.Students-per-ensenyament-i-unitats-dels/xvme-26kg]
* The database of daily COVID-19 cases in the general population by BHA, from the health department [https://dadescovid.cat/descarregues], more precisely [https://analisi.transparenciacatalunya.cat/ca/Salut/Registre-de-casos-de-COVID-19-realitzats-a-Catalun/xuwf-dxjd]
* The database of BHA and population of each area. Registre central de població del CatSalutSalut [https://analisi.transparenciacatalunya.cat/en/Salut/Registre-central-de-poblaci-del-CatSalut/ftq4-h9vk]



### Summary of Included Schools

This is the summary of Primary Schools
```{r}
kable(t(summary(primary.schools)),digits = 1,format="html",caption = "Summary of Primary Schools")
``` 
```{r}
kable(t(summary(secondary.schools)),digits = 1,format="html",caption = "Summary of Secondary Schools")
```


```{r}
#generate the totals
num.centres<-xtabs(~Type.of.School,total.centres)
num.students<-xtabs(Enrolled.Students~Type.of.School,total.centres)

# Weeks 1 and 2
mean.incidence.ABS.1.14<-cbind(mean(primary.schools$BHA.Incidence.Weeks.1.and.2,na.rm=TRUE),mean(secondary.schools$BHA.Incidence.Weeks.1.and.2,na.rm=TRUE))
sd.incidence.ABS.1.14<-cbind(sd(primary.schools$BHA.Incidence.Weeks.1.and.2,na.rm=TRUE),sd(secondary.schools$BHA.Incidence.Weeks.1.and.2,na.rm=TRUE))
#We put them together
num.of.positive.1.14<-xtabs(TOTAL.Weeks.1.and.2~Type.of.School,total.centres)
num.of.positive.students.1.14<-xtabs(Students.Weeks.1.and.2~Type.of.School,total.centres)
count.of.positive.students.1.14<-xtabs(~TOTAL.Weeks.1.and.2+Type.of.School,total.centres)

num.of.centres.with.positives.1.14<-xtabs(Exposed~Type.of.School,total.centres)
per.of.centres.with.positives.1.14<-100*num.of.centres.with.positives.1.14/num.centres

#Weeks 4 and 5
num.of.positive.22.35<-xtabs(TOTAL.Weeks.4.and.5~Type.of.School,total.centres)
num.of.positive.students.22.35<-xtabs(Students.Weeks.4.and.5~Type.of.School,total.centres)
count.of.positive.students.22.35<-xtabs(~TOTAL.Weeks.4.and.5+Type.of.School,total.centres)

total.centres$Cas <- total.centres$TOTAL.Weeks.4.and.5>0
num.of.centres.with.positives.22.35<-xtabs(Cas~Type.of.School,total.centres)
per.of.centres.with.positives.22.35<-100*num.of.centres.with.positives.22.35/num.centres


taula<-rbind(num.centres,
             num.students,
             mean.incidence.ABS.1.14,
             sd.incidence.ABS.1.14,
             num.of.positive.1.14,
             num.of.positive.students.1.14,
             count.of.positive.students.1.14,
             num.of.centres.with.positives.1.14,
             per.of.centres.with.positives.1.14,
             num.of.positive.22.35,
             num.of.positive.students.22.35,
             count.of.positive.students.22.35,
             num.of.centres.with.positives.22.35,
             per.of.centres.with.positives.22.35)

#We change the row names
noms.taula<-c("Number of Schools",
                   "Number of Enrolled Students",
                   "BHA incidence (days 0-13): mean",
                   "BHA incidence (days 0-13): sd ",
                   "Total Number of positives (days 1-14)",
                   "........of which students (days 1-14)",
                   paste(paste("Centers with", levels(factor(total.centres$TOTAL.Weeks.1.and.2)))," cases on days 1-14"),
                   "Centres with Some positive (days 1-14): Count",
                   "Centres with Some positive (days 1-14):     %",
                   "Total Number of positives (days 22-25)",
                   "........of which students (days 22-35)",
                   paste(paste("Centers with", levels(factor(total.centres$TOTAL.Weeks.4.and.5)))," cases on days 22-35"),
                   "Centres with Some positive (days 22-35): Count",
                   "Centres with Some positive (days 22-35):     %"
                   )
rownames(taula)<-noms.taula
```


```{r}
kable(taula,digits = 1, caption = 'Table 1',col.names = c('Primary Schools','Secondary Schools'))
```
## Incidence in the Basic Health Areas (ABS/BHA)

Incidence in the Basic Health Area in Weeks 1-2 (cases per 100,000) is recorded in variable 'BHA.Incidence.Weeks.1.and.2':
```{r}
incidencies<-total.centres$BHA.Incidence.Weeks.1.and.2[match(unique(total.centres$BHA),total.centres$BHA)]
summary(incidencies)
```

## Confined Students and centres

```{r}
taula.confinats<-xtabs(cbind(Confined.Students.Week.2,
            Confined.Students.Week.3,
            Confined.Students.Week.5)~Type.of.School,total.centres)
kable(taula.confinats,caption= 'Number of Confined Students', 
      col.names = c('End of Week 2','End of Week 3', 'End of Week 5'))
```
Number of confined students at the end of Week 2:
```{r}
estudiants.primaria<-sum(primary.schools$Enrolled.Students)
estudiants.secundaria<-sum(secondary.schools$Enrolled.Students)
kable(binom.exact(taula.confinats[c(1:2),1],c(estudiants.primaria,estudiants.secundaria)),caption='Confined Students. End of Week 2')
```
Number of confined students at the beginning of Week 4:
```{r}
kable(binom.exact(taula.confinats[c(1:2),2],c(estudiants.primaria,estudiants.secundaria)),caption='Confined Students. End of Week 3')
```
Number of confined students at the end of Week 5:
```{r}
kable(binom.exact(taula.confinats[c(1:2),3],c(estudiants.primaria,estudiants.secundaria)),caption='Confined Students. End of Week 5')
```

## Incidence in the centres

Incidence in Schools in Weeks 1-2:
```{r}
summary(1e5*(primary.schools$Students.Weeks.1.and.2/(primary.schools$Enrolled.Students)))
```

Incidence in Schools in the second period
```{r}
summary(1e5*(primary.schools$Students.Weeks.4.and.5/(primary.schools$Enrolled.Students- primary.schools$Students.Weeks.1.2.and.3)))
```

Incidence in Institutes in the first period
```{r}
summary(1e5*(secondary.schools$Students.Weeks.1.and.2/(secondary.schools$Enrolled.Students)))
```

Incidence in Institutes in the second period
```{r}
summary(1e5*(secondary.schools$Students.Weeks.4.and.5/(secondary.schools$Enrolled.Students- secondary.schools$Students.Weeks.1.2.and.3)))
```

Now the incidences
```{r}
1e5*binom.exact(sum(primary.schools$Students.Weeks.1.and.2),sum(primary.schools$Enrolled.Students))
1e5*binom.exact(sum(secondary.schools$Students.Weeks.1.and.2),sum(secondary.schools$Enrolled.Students))
```

Weeks 4 and 5
```{r}
1e5*binom.exact(sum(primary.schools$Students.Weeks.4.and.5),sum(primary.schools$Enrolled.Students- primary.schools$Students.Weeks.1.2.and.3))
1e5*binom.exact(sum(secondary.schools$Students.Weeks.4.and.5),sum(secondary.schools$Enrolled.Students- secondary.schools$Students.Weeks.1.2.and.3))
```


## Educational Levels in each Type of School

```{r}
addmargins(table(primary.schools$Educational.Levels))
```

```{r}
addmargins(table(secondary.schools$Educational.Levels))
```


## Graphics

We define the cuts according to Harvard levels for safe reopening (see paper for description):
```{r}
total.centres$ABS.harvard.ALL<-cut(total.centres$BHA.Incidence.Weeks.1.and.2,breaks = c(0,14*1,14*10,14*25,1200),
                                include.lowest = TRUE
                              # ,labels=c("green","yellow","orange","red")
                               )
semafor <- c("green","yellow", "orange", "red", "purple")
kable(addmargins(xtabs(~Type.of.School+ABS.harvard.ALL,total.centres)),caption='Number of Schools in BHA according to Harvard Risk Levels')
```
and now the plot
```{r}
ggplot(total.centres,aes(ABS.harvard.ALL))+
  geom_bar(aes(fill=(ABS.harvard.ALL)))+
  facet_wrap(~Type.of.School)+scale_fill_manual(values=semafor)+theme_bw()
```

##Incidence of School Students in the two study periods


```{r}
t.pos.inclosos<-xtabs(cbind(Students.Weeks.1.and.2,
                            Students.Weeks.4.and.5)~
                        Type.of.School,
                      total.centres)

t.mat.inclosos<-xtabs(cbind(Enrolled.Students,
                          (Enrolled.Students-Students.Weeks.1.2.and.3))~
                        Type.of.School,
                      total.centres)

df<-as.data.frame(t.pos.inclosos)
names(df)[names(df) == "Freq"] <- "Cases"
names(df)[names(df) == "Var2"] <- "Period"
df$Week<-df$Period
levels(df$Week)<-c("14/9-27/9","5/10-18/10")


df.Enrolled.Students<-as.data.frame(t.mat.inclosos)
names(df.Enrolled.Students)[names(df.Enrolled.Students) == "Freq"] <- "Enrolled.Students"
df$Enrolled.Students<-df.Enrolled.Students$Enrolled.Students
df$Incidence<-1e5*df$Cases/df$Enrolled.Students
dfaux<-binom.exact(df$Cases,floor(df$Enrolled.Students))
df$Incidence.lwr<-1e5*dfaux$lower
df$Incidence.upr<-1e5*dfaux$upper
kable(df[,c(1,3,4,5,6,7,8)],digits=2,col.names=c("Type of School","Cases among Students","Week",
                                                 "Enrolled Students","Incidence (cases x100,000)", "Lower CI", "Upper CI"))
```

## Figure 2: Incidence by Risk Levels in schools with and without cases in weeks 1-2
```{r}
t.pos.inclosos<-xtabs(cbind(Students.Weeks.1.and.2,
                            Students.Weeks.4.and.5)~
                        Exposed+ABS.harvard.ALL,
                      total.centres)

t.mat.inclosos<-xtabs(cbind(Enrolled.Students,
                          (Enrolled.Students-Students.Weeks.1.2.and.3))~
                        Exposed+ABS.harvard.ALL,
                      total.centres)

df<-as.data.frame(t.pos.inclosos)
names(df)[names(df) == "Freq"] <- "Cases"
names(df)[names(df) == "Var3"] <- "Period"
df$Week<-df$Period
levels(df$Week)<-c("14/9-27/9","5/10-18/10")


df.Enrolled.Students<-as.data.frame(t.mat.inclosos)
names(df.Enrolled.Students)[names(df.Enrolled.Students) == "Freq"] <- "Enrolled.Students"
df$Enrolled.Students<-df.Enrolled.Students$Enrolled.Students
df$Incidence<-1e5*df$Cases/df$Enrolled.Students
library(epitools)
dfaux<-binom.exact(df$Cases,df$Enrolled.Students)
df$Incidence.lwr<-1e5*dfaux$lower
df$Incidence.upr<-1e5*dfaux$upper
df$ABS.harvard.ALL.noms<-df$ABS.harvard.ALL
levels(df$ABS.harvard.ALL.noms)<-c("Green","Yellow","Orange","Red")
```
here is the plot
```{r}
levels(df$Exposed)<-c("No","Yes")
ggplot(subset(df,df$ABS.harvard.ALL.noms != "Green"))+
   geom_line(aes(x=Week,y=Incidence,color=Exposed,group=Exposed))+
  geom_pointrange(aes(x=Week,y=Incidence,ymin=Incidence.lwr,ymax=Incidence.upr,color=Exposed))+
  geom_errorbar(aes(x=Week,ymin=Incidence.lwr,ymax=Incidence.upr,color=Exposed),width=0.2)+
  facet_grid(cols=vars(ABS.harvard.ALL.noms),scales='free_y')+
  #ggtitle('All Public Schools 3-16+ of Catalonia')+
  labs(x="Period", y="Incidence in Schools (cases per 100,000 students)", color="Cases in the \n first 14 days?")+
  theme_bw()
```
and the data:

```{r}
kable(df[,c(1,4,5,6,7,8,9,10)],
      digits=2,
      col.names=c("Cases in Weeks 1-2?","Cases among Students","Period",
                                                 "Enrolled Students","Incidence (cases x100,000)", "Lower CI", "Upper CI", "Risk Level"))
```

# Table 2: Univariate and GLMM


First the univariate models
```{r}
total.univ.1<-glm(Students.Weeks.4.and.5~TOTAL.Weeks.1.and.2+offset(log(Enrolled.Students-Students.Weeks.1.2.and.3)),total.centres,family=poisson(link = "log"))

total.univ.2<-glm(Students.Weeks.4.and.5~Students.Weeks.1.and.2+offset(log(Enrolled.Students-Students.Weeks.1.2.and.3)),total.centres,family=poisson(link = "log"))

total.univ.3<-glm(Students.Weeks.4.and.5~Adults.Weeks.1.and.2+offset(log(Enrolled.Students-Students.Weeks.1.2.and.3)),total.centres,family=poisson(link = "log"))

total.univ.4<-glm(Students.Weeks.4.and.5~BHA.Incidence.Weeks.1.and.2.x1000+offset(log(Enrolled.Students-Students.Weeks.1.2.and.3)),total.centres,family=poisson(link = "log"))

total.univ.5<-glm(Students.Weeks.4.and.5~Type.of.School+offset(log(Enrolled.Students-Students.Weeks.1.2.and.3)),total.centres,family=poisson(link = "log"))
```

Now the Poisson multi-level:
```{r}
total.pois.re.3<-glmer(Students.Weeks.4.and.5~TOTAL.Weeks.1.and.2+BHA.Incidence.Weeks.1.and.2.x1000+Type.of.School+
                         (1|BHA)+(1|SR)+offset(log(Enrolled.Students-Students.Weeks.1.2.and.3)),total.centres,family=poisson(link = "log"))

total.pois.re.4<-glmer(Students.Weeks.4.and.5~Students.Weeks.1.and.2+
                         +Adults.Weeks.1.and.2+BHA.Incidence.Weeks.1.and.2.x1000+Type.of.School+
                         (1|BHA)+(1|SR)+offset(log(Enrolled.Students-Students.Weeks.1.2.and.3)),total.centres,family=poisson(link = "log"))
```

Table 2:
```{r}
pl <- c(
  `(Intercept)` = "(Intercept)",
  Students.Weeks.4.and.5 = "Incidence among students in the last two weeks ",
  TOTAL.Weeks.1.and.2= "Reported Cases among students and staff in the first two weeks",
  Students.Weeks.1.and.2= "Reported Cases among students in the first two weeks",
  Adults.Weeks.1.and.2="Reported Cases among staff in the first two weeks",
  BHA.Incidence.Weeks.1.and.2.x1000="Incidence in the Basic Health Area (cases per 1,000 habitants per 14 days) in the first two weeks",
  Type.of.School = "Type of School (ref. Primary)"
  )

tab_model(total.univ.1,total.univ.2,total.univ.3,total.univ.4,total.univ.5,total.pois.re.3,total.pois.re.4,pred.labels = pl, dv.labels = 'Relative Risk of Covid-19 Cases Cases on the last two weeks',digits = 3,collapse.ci = TRUE,show.p=TRUE)
```

## Supplementary Model: Exposition as categorical variable


```{r}
total.centres$exposure<-total.centres$TOTAL.Weeks.1.and.2>0
total.centres$exposure.to.students<-total.centres$Students.Weeks.1.and.2>0
total.centres$exposure.to.staff<-total.centres$Adults.Weeks.1.and.2>0

total.cate.3<-glmer(Students.Weeks.4.and.5~exposure+ 
                      BHA.Incidence.Weeks.1.and.2.x1000+Type.of.School+offset(log(Enrolled.Students-Students.Weeks.1.2.and.3))+
                   (1|BHA)+(1|SR),
                 total.centres,family=poisson(link = "log"))

total.cate.4<-glmer(Students.Weeks.4.and.5~exposure.to.staff+exposure.to.students+                    +BHA.Incidence.Weeks.1.and.2.x1000+Type.of.School+offset(log(Enrolled.Students-Students.Weeks.1.2.and.3))+
                   (1|BHA)+(1|SR),
                 total.centres,family=poisson(link = "log"))

tab_model(total.cate.3,total.cate.4, pred.labels = pl, dv.labels = 'Relative Risk of Covid-19 Cases Cases on the last two weeks',digits = 3,collapse.ci = TRUE, show.p=TRUE)
```
